G\"odel's God in Isabelle/HOL

### 2013-10-30 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex
new Isabelle/HOL version of the proof
updates for FU talk

 The formalization presented here has been carried and formally verified in the Isabelle/HOL 
 proof assistant; for more information on this system see the textbook by Nipkow, 
 Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
 at the Isabelle homepage: \url{http://isabelle.in.tum.de}.

@@

    Coq \cite{Coq} formalizations available at

@@

    with Sledgehammer \cite{Sledgehammer} performing remote calls to the higher-order automated
    theorem prover LEO-II \cite{LEO-II}. These calls then suggested respective 
    Metis \cite{Metis} calls as given below. The Metis proofs are then verified in Isabelle/HOL.
 \item For consistency checking, the Nitpick model finder \cite{Nitpick} has been employed.

@@

The types \isa{i} for possible worlds (or states) and \isa{mu} for individuals 

@@

\ i\ \ \ \ \ %

@@

\ mu\ \ \ \ %

@@

Possible worlds are connected by an accessibility relation .%

@@

This type will be abbreviated in the remainder as \isa{{\isasymsigma}}%

@@

The classical connectives $\neg, \wedge, \Rightarrow$, and $\forall$
(for individuals and over sets of individuals) and $\exists$ (over individuals) are
lifted to type $\sigma$. Further connectives could be introduced analogously. \isa{definition} 
could be used instead of \isa{abbreviation}; the latter are always fully expanded/rewritten,
which is fine here, where the focus has been on proof automation, but which would lead to 
unnessarily complex proof tasks in a purely interactive session.%

@@

\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasymPhi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline

@@

\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPhi}\ m{\isasymand}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymPhi}\ w\ {\isasymand}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline

@@

\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}p\ m{\isasymRightarrow}\ q\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ p\ w\ {\isasymlongrightarrow}\ q\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline

@@

\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline

@@

\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}i{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}i\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline

@@

\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}p{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}p\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline

@@

\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ p\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline

@@

\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ p\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ p\ v{\isacharparenright}{\isachardoublequoteclose}%

@@

For the grounding of lifted formulas the meta-predicate \isa{valid} is introduced.%

@@

@@

The model finder Nitpick confirms that the axioms and definitions above are consistent. 
Unfortunately, the respective command syntax for Nitpick is not very intuitive.%

@@

\ \ \isacommand{lemma}\isamarkupfalse%
\ True\ \isacommand{nitpick}\isamarkupfalse%
\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%

@@

\ %

@@

\isacommand{oops}\isamarkupfalse%

@@

\ P\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}%
\imp P(\psi)]$ (A property necessarily implied by a positive property is positive.).%
\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}X{\isachardot}\ {\isasymPhi}\ X\ m{\isasymRightarrow}\ {\isasympsi}\ X{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
properties are possibly exemplified). T1 is proved directly by Sledghammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}\ remote{\isacharunderscore}satallax{\isacharbrackright}}. This successful attempt then suggest to 
instead try the Metis call in the line below. This Metis call generates a proof object that is 
verified in Isabelle/HOL's kernel.%
\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline
\ A{\isadigit{2}}\ A{\isadigit{1}}a\ \isacommand{by}\isamarkupfalse%
Next, the symbol \isa{G}, for "God-like", is introduced and defined 

@@

\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%

@@

Axiom \isa{A{\isadigit{3}}} is added: $P(G)$ (The property of being God-like is positive.).

@@

(Possibly, God exists.).%

@@

\ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline

@@

We add axiom \isa{A{\isadigit{4}}}: $\all \phi [P(\phi) \to \Box \; P(\phi)]$ 

@@

\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%

@@

Symbol \isa{ess}, for "Essence", is introduced and defined as 

@@

\ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%

@@

\ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}i\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline

@@

\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}mu\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%

@@

property.).%

@@

\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}i\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline

@@

\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}i\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\isanewline

@@

Finally, the consistency of our entire theory with Nitpick.%

@@

\ \ \isacommand{lemma}\isamarkupfalse%
\ True\ \isacommand{nitpick}\isamarkupfalse%
\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
\ %
\isacommand{oops}\isamarkupfalse%

@@

produce latex documents from Isabelle sources, and he pointed to many useful tricks in Isabelle.%

>>>>

 The formalization presented here has been carried and formally verified in the Isabelle/HOL 
 proof assistant; for more information on this system see the textbook by Nipkow, 
 Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
 at: \url{http://isabelle.in.tum.de}.

@@

    Coq \cite{Coq} formalizations at

@@

    with Sledgehammer \cite{Sledgehammer}, performing remote calls to the higher-order automated
    theorem prover LEO-II \cite{LEO-II}. These calls suggest the 
    Metis \cite{Metis} calls as given below. The Metis proofs are verified in Isabelle/HOL.
 \item For consistency checking, the model finder \cite{Nitpick} has been employed.

@@

The types \isa{i} for possible worlds (or states) and $\mu$ for individuals 

@@

\ i\ \ \ \ %

@@

\ {\isasymmu}\ \ \ \ %

@@

Possible worlds are connected by an accessibility relation \isa{r}.%

@@

This type will be abbreviated in the remainder as \isa{{\isasymsigma}}.%

@@

The classical connectives $\neg, \wedge, \rightarrow$, and $\forall$
(over individuals and over sets of individuals) and $\exists$ (over individuals) are
lifted to type $\sigma$. Further connectives could be introduced analogously. Definitions 
could be used instead of abbreviations.%

@@

\ mnot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}m{\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}m{\isasymnot}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasymphi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \ \isanewline

@@

\ mand\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymand}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{9}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ m{\isasymand}\ {\isasympsi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymphi}\ w\ {\isasymand}\ {\isasympsi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline

@@

\ mimplies\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}m{\isasymRightarrow}{\isachardoublequoteclose}\ {\isadigit{7}}{\isadigit{4}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ m{\isasymRightarrow}\ {\isasympsi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymphi}\ w\ {\isasymlongrightarrow}\ {\isasympsi}\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline

@@

\ mforall{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymforall}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ {\isasymPhi}\ x\ w{\isacharparenright}{\isachardoublequoteclose}\ \ \ \isanewline

@@

\ mexists{\isacharunderscore}ind\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymexists}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}\ {\isasymPhi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ {\isasymPhi}\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline

@@

\ mforall{\isacharunderscore}indset\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymPi}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymPi}\ P\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ P\ x\ w{\isacharparenright}{\isachardoublequoteclose}\isanewline

@@

\ mbox\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasymnot}\ w\ r\ v\ {\isasymor}\ {\isasymphi}\ v{\isacharparenright}{\isachardoublequoteclose}\isanewline

@@

\ mdia\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ {\isasymphi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}v{\isachardot}\ w\ r\ v\ {\isasymand}\ {\isasymphi}\ v{\isacharparenright}{\isachardoublequoteclose}%

@@

For grounding lifted formulas, the meta-predicate \isa{valid} is introduced.%

@@

\isamarkupsection{G\"odel's Ontological Argument%

@@

@@

\ P\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}%
\imp P(\psi)]$ (A property necessarily implied by a positive property is positive).%
\ \ \ \ A{\isadigit{1}}a{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ A{\isadigit{1}}b{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ m{\isasymnot}\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ A{\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isacharparenleft}P\ {\isasymPhi}\ m{\isasymand}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPhi}\ x\ m{\isasymRightarrow}\ {\isasympsi}\ x{\isacharparenright}{\isacharparenright}{\isacharparenright}\ m{\isasymRightarrow}\ P\ {\isasympsi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
properties are possibly exemplified). T1 is proved directly by Sledghammer with command \isa{sledgehammer\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}}. This successful attempt then suggests to 
instead try the Metis call in the line below. This Metis call generates a proof object 
that is verified in Isabelle/HOL's kernel.%
\ T{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \ \isanewline
\ \ \isanewline

@@

@@

\ {\isacharparenleft}metis\ A{\isadigit{1}}a\ A{\isadigit{2}}{\isacharparenright}%

@@

Next, the symbol \isa{G} for "God-like" is introduced and defined 

@@

\ G\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}G\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymPhi}\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%

@@

Axiom \isa{A{\isadigit{3}}} is added: $P(G)$ (The property of being God-like is positive).

@@

(Possibly, God exists).%

@@

\ C{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymexists}\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isanewline

@@

Axiom \isa{A{\isadigit{4}}} is added: $\all \phi [P(\phi) \to \Box \; P(\phi)]$ 

@@

\ \isakeyword{where}\ A{\isadigit{4}}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ P\ {\isasymPhi}\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}P\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%

@@

Symbol \isa{ess} for "Essence" is introduced and defined as 

@@

\ ess\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{infixr}\ {\isachardoublequoteopen}ess{\isachardoublequoteclose}\ {\isadigit{8}}{\isadigit{5}}{\isacharparenright}\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}{\isasymPhi}\ ess\ x\ {\isacharequal}\ {\isasymPhi}\ x\ m{\isasymand}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasympsi}{\isachardot}\ {\isasympsi}\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}y{\isachardot}\ {\isasymPhi}\ y\ m{\isasymRightarrow}\ {\isasympsi}\ y{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%

@@

\ T{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ G\ x\ m{\isasymRightarrow}\ G\ ess\ x{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \isanewline

@@

\ NE\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymmu}\ {\isasymRightarrow}\ {\isasymsigma}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}NE\ {\isacharequal}\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymPi}\ {\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}\ {\isasymPhi}\ ess\ x\ m{\isasymRightarrow}\ {\isasymbox}\ {\isacharparenleft}{\isasymexists}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%

@@

property).%

@@

\ T{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymbox}\ {\isacharparenleft}{\isasymexists}\ G{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isanewline

@@

\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isanewline

@@

The consistency of the entire theory is checked with Nitpick.%

@@

@@

produce latex documents from Isabelle sources, and he showed us useful tricks in Isabelle.%
